---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fdaoutlier
Outlier Detection Tools for Functional Data

<!-- badges: start -->

[![Build Status](https://travis-ci.com/otsegun/fdaoutlier.svg?token=is9BrWwNvqBGoneFkbEL&branch=master)](https://travis-ci.com/otsegun/fdaoutlier)
  [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Codecov test coverage](https://codecov.io/gh/otsegun/fdaoutlier/branch/master/graph/badge.svg)](https://codecov.io/gh/otsegun/fdaoutlier?branch=master&token=0c40801f-206f-4bb4-9b3b-5f493a3130c9)

<!-- badges: end -->
`fdaoutlier` is a collection of outlier detection tools for functional data analysis. `fdaoutlier` is still at an early experimental stage. 

## Installation

```{r echo = F, eval = F}
install.packages("fdaoutlier")
```

Since `fdaoutlier` is still at an experimental stage, you can only install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("otsegun/fdaoutlier")
```
## Example

```{r}
library(fdaoutlier)
data(sim_data1)
# MSPLOT Dai & Genton (2018)
msplot(sim_data1$data, data_depth = "random_projection")$outliers_index
sim_data1$true_outliers
```

## Methods Implemented
1. MS-Plot (Dai & Genton, 2018)
2. TVDMSS (Huang & Sun, 2019)
3. Extremal depth (Narisetty & Nair, 2016)
4. Extreme rank length depth (Myllymäki et al., 2017; Dai et al., 2020)
5. Directional quantile (Myllymäki et al., 2017; Dai et al., 2020)
6. Fast band depth and modified band depth (Sun et al., 2012)
7. Directional Outlyingness (Dai & Genton, 2019)
8. Sequential transformation (Dai et al., 2020)

