---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fdalite
Fast Outlier Detection and Visualization Tools for Functional Data

<!-- badges: start -->
[![Build Status](https://travis-ci.com/otsegun/fdalite.svg?token=is9BrWwNvqBGoneFkbEL&branch=master)](https://travis-ci.com/otsegun/fdalite)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Codecov test coverage](https://codecov.io/gh/otsegun/fdalite/branch/master/graph/badge.svg)](https://codecov.io/gh/otsegun/fdalite?branch=master)
<!-- badges: end -->

`fdalite` is a collection of outlier detection and visualization tools for functional data analysis. `fdalite` is still at an early experimental stage. 

## Installation


```{r echo = F, eval = F}
install.packages("fdalite")
```

Since fdalite is still at an experimental stage, you can only install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("otsegun/fdalite")
```
## Example

```{r}
library(fdalite)
data(sim_data1)
# MSPLOT Dai & Genton (2018)
msplot(sim_data1$data, data_depth = "random_projection")$outliers_index

sim_data1$true_outliers
```

