---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fdaoutlier
Outlier Detection Tools for Functional Data Analysis

<!-- badges: start -->

[![Build Status](https://travis-ci.com/otsegun/fdaoutlier.svg?token=is9BrWwNvqBGoneFkbEL&branch=master)](https://travis-ci.com/otsegun/fdaoutlier)
[![Codecov test coverage](https://codecov.io/gh/otsegun/fdaoutlier/branch/master/graph/badge.svg)](https://codecov.io/gh/otsegun/fdaoutlier?branch=master&token=0c40801f-206f-4bb4-9b3b-5f493a3130c9)
[![Lifecycle: experimental](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![CRAN status](https://www.r-pkg.org/badges/version/fdaoutlier)](https://CRAN.R-project.org/package=fdaoutlier)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/fdaoutlier)](https://cran.r-project.org/package=fdaoutlier)
[![Licence](https://img.shields.io/badge/licence-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
<!-- badges: end -->
`fdaoutlier` is a collection of outlier detection tools for functional data analysis. Methods implemented include directional outlyingness, MS-plot, total variation depth, and sequential transformations among others. 

## Installation
You can install the current version of fdaoutliers from CRAN with:

```{r echo = T, eval = F}
install.packages("fdaoutlier")
```

or the latest the development version from [GitHub](https://github.com/) with:

```{r echo = T, eval = F}
devtools::install_github("otsegun/fdaoutlier")
```

## Example
Generate some functional data with magnitude outliers:

```{r, fig.align='center', fig.width = 10, fig.height=8}
library(fdaoutlier)
simdata <- simulation_model1(plot = T, seed = 1)
dim(simdata$data)
```

Next apply the msplot of Dai & Genton (2018)

```{r, fig.align='center', fig.width = 10, fig.height=8}
ms <- msplot(simdata$data)
ms$outliers
simdata$true_outliers
```



## Methods Implemented
1. MS-Plot (Dai & Genton, 2018)
2. TVDMSS (Huang & Sun, 2019)
3. Extremal depth (Narisetty & Nair, 2016)
4. Extreme rank length depth (Myllymäki et al., 2017; Dai et al., 2020)
5. Directional quantile (Myllymäki et al., 2017; Dai et al., 2020)
6. Fast band depth and modified band depth (Sun et al., 2012)
7. Directional Outlyingness (Dai & Genton, 2019)
8. Sequential transformation (Dai et al., 2020)

## Bugs and Feature Requests
Kindly open an issue using [Github issues](https://github.com/otsegun/fdaoutlier/issues).
